---
title: "JOUR405: Statistics for Journalists"
subtitle: "Midterm Exam"
name: "Ash Newton"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

## Load the tidyverse

```{r}
library(tidyverse)
```

---

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible, the compliance score, and the grade.

### Task 1: Read the data (2 points)

```{r}
scores <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```

### Task 2: Calculate the mean, standard deviation, minimum, and maximum of compliance scores. Save the results in a new dataframe called `compliance_stats`. (2 points for each)

```{r}
compliance_stats <- scores |>
  summarize(mean(scores$compliance_score), sd(scores$compliance_score), min(scores$compliance_score), max(scores$compliance_score))

```

### Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a descriptive title and label the x and y axes using `labs()`. (10 points)

```{r}
ggplot(data = scores, aes(x = compliance_score)) + 
  geom_histogram() +
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Distribution of Restaurant Health Code Compliance Scores",
    x = "Compliance Score",
    y = "Count",
  ) 
```

### Task 4: Write 2–3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What is newsworthy, and what would your next reporting step be? (5 points)

Most restaurants have high compliance scores, if not perfect, since the standard deviation is about 6. The histogram supports this interpretation and shows that an overwhelming majority of restaurants have a score of 100. Newsworthy subjects include the outliers, like that lowest score of 42 for Kenny's Sub Shop in Wheaton, MD -- I don't live far from there, I walk past that place all the time! What could they be doing to earn a score as low as 42? I would likely just walk in and order a sub, if they're even open. I would anticipate the owners would be resistant to a reporter exposing just how low their health rating is, so instead of speaking to owners or employees outright, I'd try and scope out the place without calling much attention, and then speak to health inspectors.

---

## Part 2: Daily Transit Ridership — Z-Scores and Sampling (30 points)

Your editor wants you to identify unusual days in WMATA's daily ridership data that might be worth investigating as story leads. The dataset contains daily bus and rail totals. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv`

### Task 5: Load the WMATA ridership data. (2 points)

```{r}
wmatarides <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")
```

### Task 6: Using `summarize()`, calculate the mean and standard deviation for both bus and rail ridership. Save the results in a dataframe called `wmata_stats`. (5 points)

```{r}
wmatastats <- wmatarides |>
  summarize(mean(bus), mean(rail), sd(bus), sd(rail))
```

### Task 7: Using `mutate()`, add two new columns to `wmata_daily`: `rail_zscore` (the z-score for rail ridership) and `bus_zscore` (the z-score for bus ridership). Save the result as `wmata_zscores`. (8 points)

```{r}
wmata_zscores <- wmatarides %>%
  mutate(
    mean_bus = mean(bus),
    sd_bus = sd(bus),
    bus_zscore = (bus - mean_bus) / sd_bus,
    mean_rail = mean(rail),
    sd_rail = sd(rail),
    rail_zscore = (rail - mean_rail) / sd_rail
  )
```

### Task 8: Filter `wmata_zscores` to show only days where the absolute value of `rail_zscore` is greater than 2 (unusually high or low rail ridership). Arrange by the absolute value of the z-score, with the most extreme first. How many unusual days are there? (8 points)

```{r}
unusual_days <- wmata_zscores %>%
  filter(abs(rail_zscore) > 2) %>%
  arrange(desc(abs(rail_zscore)))

```

There are 6 unusual days.

### Task 9: Using `set.seed()`, take a random sample of 91 days from `wmata_daily` (roughly 25% of the data). Calculate the mean and standard deviation for bus and rail ridership in the sample. How do the sample statistics compare to the population statistics you calculated in Task 6? Write 1–2 sentences comparing the values and explaining what this suggests about random sampling. (4 points)

```{r}
set.seed(42)

sample_91 <- wmatarides |> sample_n(91)
mean(sample_91$bus)
sd(sample_91$bus)
mean(sample_91$rail)
sd(sample_91$rail)
```
The mean and standard deviations of my sample weren't particularly far off from the overall population, but when working with figures in the tens or hundreds of thousands, it's hard to judge what constitutes a large or small margin. The sample mean for bus ridership was within about 6,000 of the population mean, which I think would indicate it's relatively close. This would suggest random sampling 91 days, about a quarter of the data, gives an accurate-ish representation of the population.

### Task 10: Write 2–3 sentences: What do the unusual days in your z-score analysis suggest as story leads? (3 points)

All my unusual days have quite low z-scores, meaning ridership was significantly lower that day than normal. I would want to investigate why people didn't take public transit on those days; were there disruptions? Was it a system problem, like a lack of service, or another factor that influences when and why people take public transit?

---

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to find out which counties have the highest rates. The dataset contains car theft counts for Maryland counties in 2023 along with population figures. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv`

### Task 11: Load the car thefts data. (2 points)

```{r}
md_car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

### Task 12: Using `mutate()`, add a column called `theft_rate` that calculates the car theft rate per 10,000 residents for each county. Save the result back to `md_car_thefts`. In 1–2 sentences, explain why a per-capita rate is more useful than raw counts for this comparison. (8 points)

```{r}
md_car_thefts <- md_car_thefts |> mutate(
  theft_rate = thefts_2023 / population *10000
)
```

A per-capita rate accounts for disparities in county population. Some counties have significantly larger populations than others, so using a per-10000 rate shows the proportion of crimes and flattens some of the larger counts.

### Task 13: Calculate the median car theft rate and the total number of car thefts statewide. Show the counties that have rates above the median. (7 points)

```{r}
median_thefts <- md_car_thefts |>
  summarize(median(theft_rate))
total_thefts <- md_car_thefts |>
  summarize(sum(thefts_2023))

median_thefts
total_thefts
```

### Task 14: Write 2–3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What is the lede of a story based on your findings? (8 points)

There were 23,871 car thefts total in Maryland in 2023. The median number of thefts was only about 10, though, which means there are certain counties with numbers high above the median skewing the overall data. What counties are those, and how many thefts did they report? What could be the cause behind the disparity?

## Part 4: Data Analysis Scenario (20 points)

A city parks department releases data showing that its 12 community swimming pools average 187 visitors per day during the summer season, with a standard deviation of 42 visitors. Your editor suspects that a handful of pools are dramatically underperforming and wants a story. You also learn that the department's figures are based on a random sample of 60 days per pool rather than a complete count.

Write 3–4 sentences for each of the following questions:

### Question A: Given that the mean is 187 and standard deviation is 42, which statistical measures would you prioritize using and why? (7 points)

A high standard of deviation means data is distributed pretty widely, so finding the interquartile range could reveal which way it skews and illustrate outliers. Alternatively, visualizing the data could easily identify outliers, such as using a histogram. If we're using a random sample of 60 days per pool rather than total attendance, I assume the figure we'd be charting is the combined attendance across those 60 days. We could also find the individual mean and standard deviation of attendance for each pool out of that sample of 60. Finding each mean and charting those would show which pools have the highest average attendance and which have the lowest -- if that sample of 60 is accurate.

### Question B: How would you use z-scores to identify which pools are genuinely underperforming, and why is that more useful than simply ranking pools by raw attendance? (7 points)

Z-scores illustrate the distance from the mean in terms of number of standard deviations. Particularly high or low z-scores demonstrate outliers that sit on the ends of the distribution. Ranking pools by raw attendance would show the highest numbers first and wouldn't account for which way the data skews.

### Question C: The department collected its attendance figures using a random sample of 60 days per pool rather than tracking every day. What questions would you ask about the sample before relying on these figures for a story? What concerns would you raise in your story? Use statistical terminology where appropriate. (6 points)

A sample of only two months out of a total year is relatively small and doesn't account for peak months during the summer vs. slow months during the winter, giving a potentially skewed portrayal of daily activity if that random sample selects lots of days from the offseason. If some pools could be underperforming, it would be better to track attendance during peak season and compare between all pools. 

---

### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your choices

Good luck!
