---
title: "HW15_GDP_DJIA"
---

What is the relationship between the stock market, represented by the Dow Jones Industrial Average, or DJIA, and the Gross Domestic Product, a measure of the U.S. economic activity?

You will examine that question for the past decade, from 2013 through 2024.

The U.S. economy in 2013 had begun to emerge from the Great Depression and heated up in the years before the COVID-19 pandemic hit. The government and private sector shutdown in a bid to save lives, ease overcrowded hospitals and prevent a wider spread of a little-known but often debilitating and deadly virus. As economic life returned to pre-pandemic levels, the outlook for the economy mostly was positive, but inflation remained high for several years.

Your task in this assignment will be to examine the relationship between those two key economic indicators, the DJIA and GDP, for the past decade. You will calculate a correlation coefficient and create a scatterplot to visualize the relationship between the two variables. As usual, you will use the Tidyverse to do this work.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the GDP and DJIA data

Load the GDP data from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/gdp_djia.csv

```{r}
gdpdjia_data <- read_csv('https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/gdp_djia.csv')
```

### Task 3: Calculate the correlation coefficient

Use the `cor()` function to calculate the correlation coefficient between the DJIA and GDP.

```{r}
gdpdjia_data |> summarize(correlation = cor(GDP, DJIA, method="pearson"))
```

### Task 4: Create a scatterplot

Use the `ggplot()` function to create a scatterplot of the DJIA and GDP. Be sure to label your axes and add a title.

```{r}
gdpdjia_data |> 
  ggplot() +
  geom_point(aes(x=GDP, y=DJIA)) +
  labs(title = "Correlation between GDP and DJIA")
```

### Task 5: Describe our results

Write a few sentences about the relationship between the DJIA and GDP. Is there a strong correlation between the two variables? What does the scatterplot show?

There is a clear positive correlation between DJIA and GDP. The correlation test brought back a number very close to 1 and the scatterplot corroborates it, showing an obvious upward trend; as GDP increases, so does DJIA.

When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.

