---
title: "Identifying Schools with Biggest Differences in Math Test Scores, Part 2"
author: "Ash Newton"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

### Introduction

In the class activity, you tested a school board member's claim using data from Montgomery County, Pa. — and found a model that didn't hold up. Now you'll run the same analysis using data from Montgomery County, Md. Your job is to evaluate whether the Maryland data tells a different story, and what that means for how you'd report on the claim.

### Loading the Data

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_md.csv")
```

### Task 1: Examine the Data First (2 points)

Before you plot anything, look at the data — you know why by now.

```{r}
summary(schools$`FARMS Pct`)
schools |> arrange(desc(`FARMS Pct`)) |> head(10)
```

Are there any data quality issues like the ones you found in the Pennsylvania data? How does the distribution of FARMS percentages compare between the two counties? Note anything that concerns you or gives you confidence in the data.

There's more data to work with than the Pennsylvania data from the activity. Well, now that I say that I feel like I'm remembering it wrong and there actually were something like 100+ schools in the PA data. There aren't any schools with a 100% FARMS percentage, unlike the PA data, and there aren't too many shockingly high math test perfomances either.


### Task 2: Create a Scatterplot and Compare (5 points)

Replace the REPLACE_ME value with the correct variable name to create a scatterplot of FARMS percentage vs. math scores for Montgomery County, Md.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations",
       title = "Math Scores vs. FARMS Percentage, Montgomery County MD")
```

Describe what you see. Then compare this scatterplot specifically to the Pennsylvania one: what differences do you notice in the direction, strength, and spread of the relationship? The two counties share a name and are both suburban — what might explain why they look so different?

The plot shows an inverse relationship between FARMS percentage and math test performance. The spread is much narrower than the PA data and the trend is stronger. One of the explanations we found in class for the PA data being so odd was a subset of outliers with 100% FARMS percentages pushing the relationship positive, but even without them, the spread was pretty random. I'm still not exactly sure why FARMS percentage didn't have much bearing on math scores in the PA data.


### Task 3: Linear Regression Model (5 points)

Replace the REPLACE_ME value with the correct variable name to run a linear regression model predicting math scores from FARMS percentage.

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)
summary(model)
```

Report the three key statistics: the p-value, the Adjusted R-squared, and the coefficient for FARMS Pct. Then interpret them together — not separately. What does the combination of these three numbers tell you about the relationship between poverty and math scores in Montgomery County, Md.? Is this model worth using? How does it compare to the Pennsylvania model?

The p-value is less than 0.00000000000000022, the adjusted r-squared is 0.6734 and the coefficient is -0.59511. Combined, this model shows that schools with higher FARMS percentage tend to perform worse on math tests. The coefficient is about -6/10, so for every 6 points lower on the math test, that FARMS percentage goes up by 10. A better way to phrase it would be in reverse, since FARMS percentage is the independent variable and math performance is the dependent, but you understand what I mean. I'm sorry, I'm like insanely burned out and slept basically the entire day and last night.


### Task 4: Identifying Schools with the Biggest Residuals (6 points)

Replace the REPLACE_ME values with the correct variable names to calculate predicted values, residuals, and absolute residuals for each school. Then arrange the schools in descending order of absolute residuals.

```{r}
schools <- schools |>
  mutate(
    predicted = predict(model),
    residual = `% Met or Exceeded Math Expectations` - predicted,
    abs_residual = abs(residual)
  ) |>
  arrange(desc(abs_residual))

head(schools |> select(1, `FARMS Pct`, `% Met or Exceeded Math Expectations`, predicted, residual), 10)
```

Identify the school with the largest positive residual (doing much better than the model predicts) and the school with the largest negative residual (doing much worse). For each school, find its zip code and look it up on [CensusReporter.org](https://censusreporter.org).

For each of the two schools, answer:
- What does the demographic and economic profile of the zip code look like?
- Does anything in that profile help explain why the school is over- or under-performing the model?
- What would your first reporting question be to the principal of that school?

Fox Chapel Elementary has the highest positive residual and is located in zipcode 20876 in Germantown, MD. The area is notable for having a 35.7% foreign-born population with most foreign-born residents originating from Asia (46%) and Latin America (38%). The median income is $119,731, slightly less than the DMV overall average of $126,684. Nothing jumps out at me that would explain why the school is overperforming the model. With a relatively well-off population, I guess I'm curious why the FARMS percentage is over 50%, but the cost of living is pretty high in MoCo, so [shrug]. My first question would be about their math program, I guess, if they have any particularly good teachers maybe.

Germantown Elementary has the highest negative residual and is located in zipcode 20874 in... Germantown, MD. Very close to Fox Chapel. Their demographics are quite similar; median income is slightly lower at $115,022. There are way more households than 20876 -- 22,662 to 8,783. Since Germantown's math performance is poor compared to their low FARMS percentage, maybe kids aren't achieving because the school is overcrowded. I would ask the principal about class sizes.


### Task 5: Evaluate the Model and Write the Pitch (7 points)

You've now run this analysis on two counties. Return to the school board member's claim: that elementary schools with fewer students on free and reduced-price meals do better on standardized math tests.

Write a short story pitch of 4-6 sentences that addresses all of the following: Does the Maryland data support or contradict the claim? How does it compare to what you found in Pennsylvania? What would you ask the school board member if you called them for comment? And what's the most interesting specific finding from your residuals analysis: which school would you want to visit first, and why?

The Maryland data supports the claim that schools with fewer students on free and reduced-price meals do better on standardized math tests. The Pennsylvania data does not support that claim, and I'm curious as to why. I would ask why some schools have FARMS percentages of 100%. The most interesting specific finding from my residuals analysis is Germantown Elementary being potentially overcrowded, due to the significantly higher number of households in the area compared to Fox Chapel. I would visit and see if that could be a factor as to why they underperform at math.
