---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2025.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

All salaries are yearly, or "per annum". Some are 0, which could cause problems for analysis, and also make me wonder why someone would work without pay -- the positions that receive 0 pay are titles like Senior Advisor and Senior Counsel to the President. Some are cabinet members like Marco Rubio. Why wouldn't this data report their salaries?

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
summary(wh_salaries)
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
summary(wh_salaries)
min(wh_salaries)
max(wh_salaries)
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The median salary is $101,500. The mean is $109,117, meaning they're largely equal and salary distribution is normal. The minimum is 0, which likely does not reflect the actual lowest salary in the data; the maximum is $225,700. I would guess, considering the positions of the people with a salary listed as 0, that their actual pay is higher than the maximum.

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

Why does this data hide the salaries of incredibly powerful White House officials? I would research into who they are and how the White House collects and publishes its data. 
