---
title: "HW8_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```
### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
mnms <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```
### Task 3: Sample 5 rows of data from the dataset
```{r}
sample_5 <- mnms |> sample_n(5)

```
### Task 4: Calculate the mean age and standard deviation of the sample, using either the blue or white number of M&Ms
```{r}
mean_mnms_sample_5 <- mean(sample_5$red)
sd_mnms_sample_5 <- sd(sample_5$red)

mean_mnms_sample_5
sd_mnms_sample_5
```
### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4
```{r}
mean_mnms_pop <- mean(mnms$red)
sd_mnms_pop <- sd(mnms$red)

mean_mnms_pop
sd_mnms_pop
```
### Task 6: Compare the sample and population statistics

Compare the two sets of means and standard deviations. Are they similar?

The means are quite similar, but the standard deviations are slightly off. The standard deviation for the entire population is higher than the standard deviation for the small sample.

### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

I think I got lucky and happened to select a sample of 5 with a mean and standard deviation that closely matches the population mean. Generally, a sample of only 5 wouldn't be enough to accurately represent the entire population.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
