---
title: "Final_Project"
output: html_document
date: "2026-04-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Giving this another pass. I was getting an NA result likely because there are NA values somewhere in Average_rating.
## Using cor.test() instead will ignore those NA values, so I can successfully test for correlation between average rating and watch count.
## I tested it using owner rating, found that it worked, then tried finding the correlation between watch count and average rating. Cor.test() also gives a p-value.

```{r cars}
library(tidyverse)
movie_data <- read_csv("~/GitHub/jour405_spring26/project/Movie_Data_File.csv")
cor.test(Movie_Data_File$Owner_rating, Movie_Data_File$Average_rating, method="pearson")
cor.test(Movie_Data_File$Watches, Movie_Data_File$Average_rating, method="pearson")
```
### 3-5 tests I can run and reasoning:
### 1. Linear regression model testing ratings against watches. That would give me an understanding of their relationship. I need to remember that all I'm looking for is whether the most-watched films on Letterboxd are the most-liked.
### 2. Histogram of ratings. Is a histogram a test? It's a graph. I should graph how ratings are distributed across this sample of over 10,000 films, so we can understand what it means for a film to be highly rated; is it all that rare? Even if highly watched films are in fact highly rated, is that particularly valuable?
### 2.5. Another test that would help understand the distribution of the data is finding the interquartile range. I could also make box plots, but I find them hard to interpret.
### 3. Correlation tests to check my linear regression models. This will help verify what the models indicate. I should also compare total ratings to average rating and number of watches. There may be a discrepancy between number of ratings and number of watches that could be affecting the data. Like, I don't want to posit that the most-watched films on Letterboxd are also the most well-liked, if they're actually not being rated by very many people. I'm giving too much credit to that minority.