---
title: "Final_Project"
output: html_document
date: "2026-04-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
movie_data <- read_csv("~/GitHub/jour405_spring26/project/Movie_Data_File.csv")
```

### For this project, I'm analyzing a population of 10,000+ films on the review aggregator platform Letterboxd. I want to understand if the most-watched films are the highest rated. Here's a series of tests to understand the relationship between these variables.

### Correlation test:

```{r}
cor.test(x=log10(movie_data$Watches), y=movie_data$Average_rating, method="pearson")
```
### The correlation is 0.27, or slightly positive. Overall, there is a weak positive correlation between watch count and rating. The p-value is incredibly small due to the massive sample size, but demonstrates that this is a statistically significant finding. We use the log(10) of the total watch count to try and cut down the influence of outliers on the far end.

### Summary statistics:

```{r}
summary(movie_data$Average_rating)
summary(movie_data$Watches)
```
### The overall average rating is 3.26. The rating scale on Letterboxd is a ten-point scale allowing ratings from 0.5 to 5. There's no midpoint rating; a rating of 3 stars falls on the upper half of the rating scale. Ratings could be said to trend slightly positive. 

### The average watch count is significantly higher than the median, meaning it's being influenced by some particularly high outliers. Being higher than even the third quarter, it's likely that the maximum/highest data point of over 5 million watches for a single film is to blame. The median is less than 20,000, which gives a more realistic understanding of watch count distribution. 

### To elaborate on distribution, here are some histograms.

```{r}
ggplot(movie_data, aes(x=Watches)) +
  geom_histogram() +
  labs(x="Watches", y="Count") +
  scale_x_log10(labels = scales::label_comma())

ggplot(movie_data, aes(x=Average_rating)) +
  geom_histogram() +
  labs(x="Average Rating", y="Count")
```

### With this information, we can now compare these two distributions. Linear regression of watch count versus average rating:

```{r}
model <- lm(Average_rating ~ log10(Watches), data = movie_data)
summary(model)
```

### An adjusted r-squared of 0.7 illustrates that very little of the variance in average rating is determined by watch count. Let's visualize this model:

```{r}
movie_data |>
  ggplot(aes(x = Watches, y = Average_rating)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    scale_x_log10(labels = scales::label_comma()) +
    labs(x = "Watches (log scale)", y = "Average Rating")
```

### Visually, it's easy to tell the mild positive correlation between watches and average rating, and the incredibly wide variance in ratings. Notably, the data also tapers on the ends; points with relatively few watches cluster around a 3 to 3.5 average rating. As watch count increases, the variance does as well, dipping down to ratings near 1 star before tapering back up again.

### Let's take a closer look at the films with a high watch count. Mean rating of top 100 watched films versus total:

```{r}
top_watched <- movie_data |>
  arrange(desc(Watches)) |>
  slice_head(n = 100)

summary(top_watched$Average_rating)
mean(top_watched$Average_rating, na.rm = TRUE)
mean(movie_data$Average_rating, na.rm = TRUE)
```
### The average rating for the top 100 most watched films in the dataset is slightly higher than the overall average.

### All of this analysis combined leads me to believe that this data isn't particularly conducive to a story. While there is a mild relationship between watch count and average rating, the lukewarm correlation raises more questions than it answers. As such, if I were to prepare a story on the highest-rated films on Letterboxd, I would prefer to investigate if there are any commonalities between them. What kinds of films do Letterboxd users like? Are there aspects of a film that could help predict how it will perform critically before it's even released? This dataset does include information on director, cast, runtime, genre and language for each film, so an analysis of this kind is feasible.